---
params:
  subtitle: ""
  title: ""
  authors: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: '`r params$authors`'
date: 'Last run at: `r getRunDateTime()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
bibliography: '`r paste0(here::here(), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```


```{r codeSetup, include=FALSE}
# Set start time ----
startTime <- proc.time()

# Local parameters ----
b2Kb <- 1024 #http://whatsabyte.com/P1/byteconverter.htm
b2Mb <- 1048576
plotLoc <- paste0(repoParams$repoLoc, "/docs/plots/") # where to put the plots (if any)

# Packages used in the report ----
rmdLibs <- c("ggplot2", # plots
          "kableExtra", # fancy tables
          "hms", # times
          "skimr", # for skim
          "tidyr" # for gather
          )
GREENGridEECA::loadLibraries(rmdLibs)

# Local functions ----



```

\newpage

# About

## Citation

```{r citation, child=repoParams$citation}
```

## Report circulation:

 * Public – this report is intended for publication following EECA approval.
 
## License

```{r ccby license, child=repoParams$licenseCCBY}
```
 
## History

```{r history, child=repoParams$history}
```
 * this [report's edit history](https://github.com/CfSOtago/GREENGridEECA/commits/master/reports/partB_dataAnalysis)
 
## Support

```{r generic support, child=repoParams$support}
```
 
\newpage

# Introduction

This report uses the GREEN Grid project [@stephenson_smart_2017] research data to analyse a variety of residential household appliances and their contribution to peak demand under several scenarios.

# Data

```{r generic sample, child=repoParams$data}
```

```{r partBdataProcessing}
# put all this here before you report on it
# this code _could_ go in the makeFile if you want...

#New DT
powerDT <- copy(origPowerDT) # need to copy as data.table works by reference https://www.rdocumentation.org/packages/data.table/versions/1.12.2/topics/copy

#DT modifications

#Excluding households based on Part A report: https://cfsotago.github.io/GREENGridEECA/#part-a--data-processing
exclude <- c("rf_14", "rf_25", "rf_26", "rf_43", "rf_46")
powerDT <- powerDT[!(linkID %in% exclude)]

# setting negative values to NA
powerDT <- powerDT[, meanPowerW := ifelse(meanPowerW <0, NA, meanPowerW)]
powerDT <- powerDT[, sdPowerW := ifelse(sdPowerW <0, NA, sdPowerW)]
powerDT <- powerDT[, minPowerW := ifelse(minPowerW <0, NA, minPowerW)]
powerDT <- powerDT[, maxPowerW := ifelse(maxPowerW <0, NA, maxPowerW)]


# set to NZ time from UTC
powerDT <- powerDT[, r_dateTime_nz := lubridate::as_datetime(r_dateTimeHalfHour, 
                                               tz = "Pacific/Auckland")] # this will be UTC unless you set this

#Define Winter/else
powerDT <- powerDT[, date := lubridate::date(r_dateTime_nz)]
powerDT <- powerDT[, obsHalfHour := hms::as.hms(r_dateTime_nz)]
#powerDT[, obsHalfHour := format(ymd_hms(r_dateTimeHalfHour), "%H:%M:%S")]
powerDT <- powerDT[, month := lubridate::month(r_dateTime_nz)]
powerDT <- powerDT[, peak := 0]
powerDT[, year := lubridate::year(r_dateTime_nz)] # so we can do per-year analysis

powerDT <- powerDT[month == 12 | month == 1 | month == 2, season := "Summer"]
powerDT <- powerDT[month == 3 | month == 4 | month == 5, season := "Autumn"]
powerDT <- powerDT[month == 6 | month == 7 | month == 8, season := "Winter"]
powerDT <- powerDT[month == 9 | month == 10 | month == 11, season := "Spring"]

#Setting times of peak demand 
OP1S <- hms::as.hms("00:00:00")
OP1E <- hms::as.hms("16:30:00")

PS <- hms::as.hms("17:00:00")
PE <- hms::as.hms("21:00:00")

OP2S <- hms::as.hms("21:30:00")
OP2E <- hms::as.hms("23:30:00")

powerDT <- powerDT[, peak := ifelse(obsHalfHour >= OP1S & obsHalfHour <= OP1E,
                                    "Off Peak 1",
                                    NA)]
powerDT <- powerDT[, peak := ifelse(obsHalfHour >= OP2S & obsHalfHour <= OP2E,
                                    "Off Peak 2",
                                    peak)]
powerDT <- powerDT[, peak := ifelse(obsHalfHour >= PS & obsHalfHour <= PE,
                                    "Peak",
                                    peak)]
# this stops the RHS coercion errors
```

```{r dailyMeans }

#Creating daily avergaes 

dailyAvgDT <- copy(powerDT)

dailyAvgDT <- GREENGridEECA::labelEECACircuits(dailyAvgDT)

dailyAvgDT <- dailyAvgDT[, davgW := mean(meanPowerW, na.rm = TRUE), keyby =.(obsHalfHour, eecaCircuit, year, season)]

dailyAvgDT <- dailyAvgDT[, .(davgW=davgW), keyby=.(obsHalfHour, eecaCircuit, year, season)]
dailyAvgDT <- unique(dailyAvgDT)# Removing duplicates
dailyAvgDT$season <- factor(dailyAvgDT$season, levels = c("Spring","Summer",
                                                    "Autumn", "Winter"))#This puts the seasons in the right order in the plot
write.csv(dailyAvgDT, file = "GG_daily_average_power_multiple_years.csv")
```

```{r LighringPlot}

myPlot <- ggplot2::ggplot(dailyAvgDT[eecaCircuit =="Lighting" &year =="2015"], aes(x=obsHalfHour, colour=season)) +
  geom_line(aes(y=davgW))+
  facet_grid(season ~ .) +
  labs(x='Time of Day', y='Power in W') +
  scale_x_time(breaks = c(hms::as.hms("00:00:00"), 
                          hms::as.hms("04:00:00"), 
                          hms::as.hms("08:00:00"),       
                          hms::as.hms("12:00:00"), 
                          hms::as.hms("16:00:00"), 
                          hms::as.hms("20:00:00"),
                          hms::as.hms("24:00:00")))+
  
  theme(text = element_text(family = "Cambria"),
        axis.text.y = element_text(colour = "black"),
        axis.text.x = element_text(colour = "black"))+
  ggtitle("GG sample daily average lighting demand by season for 2015 in Watts")
myPlot

#ggsave("GG sample daily average lighting demand by season for 2015 in Watts.jpeg", dpi = 900)

```
```{r HotWaterPlot}

myPlot <- ggplot2::ggplot(dailyAvgDT[eecaCircuit =="Hot water" &year =="2015"], aes(x=obsHalfHour, colour=season)) +
  geom_line(aes(y=davgW))+
  facet_grid(season ~ .) +
  labs(x='Time of Day', y='Power in W') +
  scale_x_time(breaks = c(hms::as.hms("00:00:00"), 
                          hms::as.hms("04:00:00"), 
                          hms::as.hms("08:00:00"),       
                          hms::as.hms("12:00:00"), 
                          hms::as.hms("16:00:00"), 
                          hms::as.hms("20:00:00"),
                          hms::as.hms("24:00:00")))+
  
  theme(text = element_text(family = "Cambria"),
        axis.text.y = element_text(colour = "black"),
        axis.text.x = element_text(colour = "black"))+
  ggtitle("GG sample daily average hot water demand by season for 2015 in Watts")
myPlot

#ggsave("GG sample daily average hot water demand by season for 2015 in Watts.jpeg", dpi = 900)

```
```{r OvenPlot}

myPlot <- ggplot2::ggplot(dailyAvgDT[eecaCircuit =="Oven" &year =="2015"], aes(x=obsHalfHour, colour=season)) +
  geom_line(aes(y=davgW))+
  facet_grid(season ~ .) +
  labs(x='Time of Day', y='Power in W') +
  scale_x_time(breaks = c(hms::as.hms("00:00:00"), 
                          hms::as.hms("04:00:00"), 
                          hms::as.hms("08:00:00"),       
                          hms::as.hms("12:00:00"), 
                          hms::as.hms("16:00:00"), 
                          hms::as.hms("20:00:00"),
                          hms::as.hms("24:00:00")))+
  
  theme(text = element_text(family = "Cambria"),
        axis.text.y = element_text(colour = "black"),
        axis.text.x = element_text(colour = "black"))+
  ggtitle("GG sample daily average oven demand by season for 2015 in Watts")
myPlot

#ggsave("GG sample daily average oven demand by season for 2015 in Watts.jpeg", dpi = 900)

```
```{r heatpumpPlot}

myPlot <- ggplot2::ggplot(dailyAvgDT[eecaCircuit =="Heat Pump or Heating" &year =="2015"], aes(x=obsHalfHour, colour=season)) +
  geom_line(aes(y=davgW))+
  facet_grid(season ~ .) +
  labs(x='Time of Day', y='Power in W') +
  scale_x_time(breaks = c(hms::as.hms("00:00:00"), 
                          hms::as.hms("04:00:00"), 
                          hms::as.hms("08:00:00"),       
                          hms::as.hms("12:00:00"), 
                          hms::as.hms("16:00:00"), 
                          hms::as.hms("20:00:00"),
                          hms::as.hms("24:00:00")))+
  
  theme(text = element_text(family = "Cambria"),
        axis.text.y = element_text(colour = "black"),
        axis.text.x = element_text(colour = "black"))+
  ggtitle("GG sample daily average heat pump demand by season for 2015 in Watts")
myPlot

#ggsave("GG sample daily average heat pump demand by season for 2015 in Watts.jpeg", dpi = 900)

```