---
params:
  subtitle: ""
  title: ""
  authors: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: '`r params$authors`'
date: 'Last run at: `r getRunDateTime()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
    fig_width: 5
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
    fig_width: 5
always_allow_html: yes
bibliography: '`r paste0(here::here(), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```


```{r codeSetup, include=FALSE}
# Set start time ----
startTime <- proc.time()
# Local parameters ----
b2Kb <- 1024 #http://whatsabyte.com/P1/byteconverter.htm
b2Mb <- 1048576
plotLoc <- paste0(repoParams$repoLoc, "/docs/plots/") # where to put the plots

# Packages used in the report ----
rmdLibs <- c("ggplot2", # plots
          "kableExtra" # fancy tables
          )
GREENGridEECA::loadLibraries(rmdLibs)
# Local functions ----
```

\newpage

# About

## Citation

```{r citation, child=repoParams$citation}
```

## Report circulation

 * Public – this report is intended for publication following EECA approval.
 
## License {#license}

```{r ccby license, child=repoParams$licenseCCBY}
```
 
## History

```{r history, child=repoParams$history}
```
 * this [report's edit history](https://github.com/CfSOtago/GREENGridEECA/commits/master/reports/partA_dataProcessing)
 
## Support

```{r generic support, child=repoParams$support}
```
 
\newpage

# Introduction

Previous work conducted as part of this project has calculated the % contribution to peak power load (kW) and to overall annual consumption (kWh) for the ~ 40 New Zealand GREEN Grid household electricity demand study dwellings [@stephenson_smart_2017].

EECA have expressed an interest in understanding how these results and/or this sample of dwellings can be scaled up to provide estimates of the same values for the overall New Zealand residential dwelling stock/household population.


As we have noted the sample of approximately 40 households used to do this is not likely to be representative of the 

# Summary

This report XXXX

This report described the results of this work and directed the reader to relevant R code where necessary.

All code used to create this report is available from:

 * https://github.com/CfSOtago/GREENGridEECA
 
The archived and most recent version of the report is available from:
 
 * https://cfsotago.github.io/GREENGridEECA/


```{r runToEnd}

```

# Data Annex

## Census data summary 
Descriptive statistics for census data:

```{r skimData}
skimr::skim(censusDT)
```

## GREEN Grid half-hourly total load summary 
Descriptive statistics for dwelling power data:

```{r skimData}
skimr::skim(powerDT)
```

## Census data summary 
Descriptive statistics for household data:

```{r skimData}
skimr::skim(householdDT)
```

# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime
elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

## R packages used

 * base R [@baseR]
 * bookdown [@bookdown]
 * data.table [@data.table]
 * ggplot2 [@ggplot2]
 * kableExtra [@kableExtra]
 * knitr [@knitr]
 * lubridate [@lubridate]
 * rmarkdown [@rmarkdown]

## Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
