---
params:
  subtitle: ""
  title: ""
  version: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: '`r repoParams$Authors`'
date: 'Last run at: `r getRunDateTime()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
bibliography: '`r paste0(here::here(), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```


```{r codeSetup, include=FALSE}
# Set start time ----
startTime <- proc.time()
# Local parameters ----
b2Kb <- 1024 #http://whatsabyte.com/P1/byteconverter.htm
b2Mb <- 1048576
plotLoc <- paste0(repoParams$repoLoc, "/docs/plots/") # where to put the plots
  
# Local functions ----
```

\newpage

# About

## Report circulation:

 * Public – this report is intended for publication.
 
## License

```{r ccby license, child=repoParams$licenseCCBY}
```
 
## Citation

```{r citation, child=repoParams$citation}
```

## History

```{r history, child=repoParams$history}
```
 * this [report's edit history](https://github.com/CfSOtago/GREENGridData/commits/master/makeDocs/buildOverviewReport.Rmd)
 
## Support

```{r generic support, child=repoParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=repoParams$sampleGeneric}
```
 
This report uses the GREEN Grid project [@stephenson_smart_2017] research data to...

# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime
elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

## R packages used

 * base R [@baseR]
 * bookdown [@bookdown]
 * data.table [@data.table]
 * ggplot2 [@ggplot2]
 * kableExtra [@kableExtra]
 * knitr [@knitr]
 * lubridate [@lubridate]
 * rmarkdown [@rmarkdown]

## Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
